---
title: Renter survey
format:
  360-analysis-html: default
author: James Goldie
date: last-modified
code-fold: true
---

```{r}
#| label: setup
library(tidyverse)
library(haven)
library(glue)
library(here)
library(sf)

# load the survey data. exclude people without a postcode
survey <-
  read_dta(here("data", "02_AHCD_Sensitive_Data_File_001469.dta")) %>%
  filter(Q43 != 97, !is.na(Q43)) %>%
  mutate(Q43 = as.character(Q43))
```

Let's also merge in the NatHERS Climate Zone data so that we can see how many people are in each climate zone.

```{r}
#| label: climate-zone-merge

here("data", "nathers-zones.csv") %>%
  read_csv() %>%
  right_join(survey, by = c("postcode" = "Q43"), multiple = "all") ->
survey_merged
```

We'll also need the town names for each zone so that people can find the zones without a map:

```{r}
#| label: town-names

here("data", "nathers-zone-names.csv") %>%
  read_csv(col_types = "ic") %>%
  mutate(town = paste0(zone, " - ", town)) ->
town_list
```

Now, how many people are in each zone?

```{r}
#| label: postcode-hist
survey_merged %>%
  count(zone) %>%
  {
    ggplot(.) +
      aes(x = zone, y = n) +
      geom_col() +
      coord_flip()
  }
```

Okay, so these results are very uneven in terms of the distribution of people. In fact, `r survey_merged %>% count(zone) %>% filter(n < 10) %>% nrow()` zones have less than 10 respondants.

As well as communicating statistically significant results, we may want to be mindful about redacting responses in zones with small response counts. Or perhaps lumping those smaller zones together (although we can't make any climate-specific inferences about those).

## Responses by zone

Let's also add margin of error at 95% confidence to see which climate zones might exclude on the basis of low population:

```{r}
#| label: helper-moe
moe_95p <- function(p, n) {
  1.96 * sqrt((p * (1 - p)) / n)
}
```

And here's a barchart function to visualise our responses:

```{r}
#| label: helper-stacked-barchart
stacked_barchart <- function(df, col_prefix, title = "Title here") {

  df %>%
    select(zone, starts_with(col_prefix)) %>%
    # standardise column names
    rename(response = !!col_prefix) %>%
    rename_with(~ str_remove_all(.x, paste0(col_prefix, "_"))) %>%
    ungroup() %>%
    mutate(zone_label = glue("Zone {zone} (n = {ntotal})")) ->
  df_tidy
  
  p_base <- ggplot(df_tidy) +
    aes(x = fct_reorder(zone_label, ntotal), y = p, fill = response) +
    geom_col(position = "stack") +
    scale_y_continuous(labels = scales::label_percent()) +
    coord_flip() +
    scale_fill_brewer(type = "div", palette = "RdBu", direction = -1) +
    theme_minimal(base_size = 16) +
    labs(title = title)

  return(p_base)
    
}
```

```{r}
#| label: helper-errorbar-barchart
errorbar_barchart <- function(df, col_prefix, title = "Title here") {

  df %>%
    select(zone, starts_with(col_prefix)) %>%
    # standardise column names
    rename(response = !!col_prefix) %>%
    rename_with(~ str_remove_all(.x, paste0(col_prefix, "_"))) %>%
    ungroup() %>%
    mutate(
      zone_label = glue("Zone {zone} (n = {ntotal})"),
      zone_label = fct_reorder(zone_label, ntotal, .desc = TRUE)) ->
  df_tidy

  # print(levels(df_tidy$zone_label))

  print(df_tidy)
  
  p_base <- ggplot(df_tidy) +
    aes(x = response, y = p, fill = response) +
    geom_col() +
    geom_linerange(aes(ymin = plow, ymax = phigh)) +
    facet_wrap(vars(zone_label)) +
    scale_y_continuous(labels = scales::label_percent()) +
    coord_flip() +
    scale_fill_brewer(type = "div", palette = "RdBu", direction = -1) +
    theme_minimal(base_size = 16) +
    labs(title = title)

  return(p_base)
    
}
```

### Heat, cold and housing condition

#### Q12: Can you keep comfortably warm in your house during winter?

* Excluding "Not applicable, have not experienced living in house during winter"

```{r}
#| label: comfortable-winter
survey_merged %>%
  select(zone, Q12) %>%
  mutate(Q12 = as_factor(Q12)) %>%
  filter(Q12 %in% c("Yes", "No")) %>%
  count(zone, Q12, name = "Q12_n") %>%
  group_by(zone) %>%
  mutate(
    Q12_ntotal = sum(Q12_n),
    Q12_p = Q12_n / sum(Q12_n),
    Q12_moe = moe_95p(p = Q12_p, n = Q12_ntotal),
    Q12_plow = Q12_p - Q12_moe,
    Q12_phigh = Q12_p + Q12_moe) %>%
  arrange(zone) ->
prop_Q12

prop_Q12 %>%
  errorbar_barchart("Q12",
    "Can you keep comfortably warm in your house during winter?")
```

#### Q13: Can you keep comfortably cool in your house during summer?

* Excluding "Not applicable, have not experienced living in house during summer"

```{r}
#| label: comfortable-summer
survey_merged %>%
  select(zone, Q13) %>%
  mutate(Q13 = as_factor(Q13)) %>%
  filter(Q13 %in% c("Yes", "No")) %>%
  count(zone, Q13, name = "Q13_n") %>%
  group_by(zone) %>%
  mutate(
    Q13_ntotal = sum(Q13_n),
    Q13_p = Q13_n / sum(Q13_n),
    Q13_moe = moe_95p(p = Q13_p, n = Q13_ntotal),
    Q13_plow = Q13_p - Q13_moe,
    Q13_phigh = Q13_p + Q13_moe) %>%
  arrange(zone) ->
prop_Q13

prop_Q13 %>%
  errorbar_barchart("Q13",
    "Can you keep comfortably cool in your house during summer?")
```

#### Q18: How would you rate the overall condition of your current home, such as the walls, roof, doors and windows for example.

* Excluding "Don't know"

```{r}
#| label: condition-house
survey_merged %>%
  select(zone, Q18) %>%
  mutate(Q18 = as_factor(Q18)) %>%
  filter(Q18 != "Don't know") %>%
  count(zone, Q18, name = "Q18_n") %>%
  # fill the missing responses (due to 0 counts) back in
  complete(Q18, zone, fill = list(Q18_n = 0)) %>%
  group_by(zone) %>%
  mutate(
    Q18_ntotal = sum(Q18_n),
    Q18_p = Q18_n / sum(Q18_n)) %>%
  arrange(zone) ->
prop_Q18
```

### Rental affordability

#### Q25: What is your household rent per month?

```{r}
#| label: monthly-rent
survey_merged %>%
  select(zone, Q25) %>%
  filter(!(Q25 %in% c(99998, 99999))) %>%
  mutate(Q25 = zap_labels(Q25)) %>%
  # TODO - bin up rent for comparison with outcomes?
  {
    ggplot(.) +
      aes(x = Q25, group = zone) +
      geom_freqpoly(alpha = 0.4) +
      scale_x_log10(
        labels = scales::label_number(
          prefix = "$",
          scale_cut = scales::cut_short_scale())) +
      labs(
        x = "Rent per month",
        y = "Number of people",
        title = "Household rent per month",
        subtitle = "In each climate zone")
  }
```

#### Q26: In general, how affordable is this amount for your household?

```{r}
#| label: affordability
survey_merged %>%
  select(zone, Q26) %>%
  mutate(Q26 = as_factor(Q26)) %>%
  count(zone, Q26, name = "Q26_n") %>%
  # fill the missing responses (due to 0 counts) back in
  complete(Q26, zone, fill = list(Q26_n = 0)) %>%
  filter(Q26 != "[DNRO] Don't know") %>%
  # get proportions
  group_by(zone) %>%
  mutate(
    Q26_ntotal = sum(Q26_n),
    Q26_p = Q26_n / sum(Q26_n)) %>%
  arrange(zone) ->
prop_Q26
```

#### Q27: What impact would a 10% increase in your rent have on your current financial situation?

Looks like all the data is missing here.

```{r}
#| label: affordability
survey_merged %>% pull(Q27) %>% summary()
```

#### Q28: After paying your rent, is there typically enough money left for ...?

##### Essential spending (bills, clothing, transport, food and drink, etc.)

```{r}
#| label: affordability-essentials
survey_merged %>%
  select(zone, Q28_01) %>%
  mutate(Q28.01 = as_factor(Q28_01)) %>%
  count(zone, Q28.01, name = "Q28.01_n") %>%
  # fill the missing responses (due to 0 counts) back in
  complete(Q28.01, zone, fill = list(Q28.01_n = 0)) %>%
  filter(Q28.01 %in% c("Yes", "No")) %>%
  # get proportions
  group_by(zone) %>%
  mutate(
    Q28.01_ntotal = sum(Q28.01_n),
    Q28.01_p = Q28.01_n / sum(Q28.01_n),
    Q28.01_moe = moe_95p(p = Q28.01_p, n = Q28.01_ntotal),
    Q28.01_plow = Q28.01_p - Q28.01_moe,
    Q28.01_phigh = Q28.01_p + Q28.01_moe) %>%
  arrange(zone) ->
prop_Q28.01

prop_Q28.01 %>%
  multichoice_barchart("Q28.01",
    paste(
      "After paying your rent, is there typically enough",
      "money left for essential spending?",
      sep = "\n"))
prop_Q28.01 %>%
  errorbar_barchart("Q28.01",
    paste(
      "After paying your rent, is there typically enough",
      "money left for essential spending?",
      sep = "\n"))
```

##### Non-essential spending (social activities, holidays, TV, nonessential food and drink, and alcohol)

```{r}
#| label: affordability-nonessentials
survey_merged %>%
  select(zone, Q28_02) %>%
  mutate(Q28.02 = as_factor(Q28_02)) %>%
  count(zone, Q28.02, name = "Q28.02_n") %>%
  # fill the missing responses (due to 0 counts) back in
  complete(Q28.02, zone, fill = list(Q28.02_n = 0)) %>%
  filter(Q28.02 %in% c("Yes", "No")) %>%
  # get proportions
  group_by(zone) %>%
  mutate(
    Q28.02_ntotal = sum(Q28.02_n),
    Q28.02_p = Q28.02_n / sum(Q28.02_n),
    Q28.02_moe = moe_95p(p = Q28.02_p, n = Q28.02_ntotal),
    Q28.02_plow = Q28.02_p - Q28.02_moe,
    Q28.02_phigh = Q28.02_p + Q28.02_moe) %>%
  arrange(zone) ->
prop_Q28.02

prop_Q28.02 %>%
  multichoice_barchart("Q28.02",
    paste(
      "After paying your rent, is there typically enough",
      "money left for non-essential spending?",
      sep = "\n"))
prop_Q28.02 %>%
  errorbar_barchart("Q28.02",
    paste(
      "After paying your rent, is there typically enough",
      "money left for non-essential spending?",
      sep = "\n"))
```

##### Savings or investment

```{r}
#| label: affordability-savings
survey_merged %>%
  select(zone, Q28_03) %>%
  mutate(Q28.03 = as_factor(Q28_03)) %>%
  count(zone, Q28.03, name = "Q28.03_n") %>%
  # fill the missing responses (due to 0 counts) back in
  complete(Q28.03, zone, fill = list(Q28.03_n = 0)) %>%
  filter(Q28.03 %in% c("Yes", "No")) %>%
  # get proportions
  group_by(zone) %>%
  mutate(
    Q28.03_ntotal = sum(Q28.03_n),
    Q28.03_p = Q28.03_n / sum(Q28.03_n),
    Q28.03_moe = moe_95p(p = Q28.03_p, n = Q28.03_ntotal),
    Q28.03_plow = Q28.03_p - Q28.03_moe,
    Q28.03_phigh = Q28.03_p + Q28.03_moe) %>%
  arrange(zone) ->
prop_Q28.03

prop_Q28.03 %>%
  stacked_barchart("Q28.03",
    paste(
      "After paying your rent, is there typically enough",
      "money left for savings or investment?",
      sep = "\n"))
prop_Q28.03 %>%
  errorbar_barchart("Q28.03",
    paste(
      "After paying your rent, is there typically enough",
      "money left for savings or investment?",
      sep = "\n"))
``` 

### Health

#### Q29: In general, would you say your health is...?

```{r}
#| label: health-general
survey_merged %>%
  select(zone, Q29) %>%
  mutate(Q29 = as_factor(Q29)) %>%
  count(zone, Q29, name = "Q29_n") %>%
  # fill the missing responses (due to 0 counts) back in
  complete(Q29, zone, fill = list(Q29_n = 0)) %>%
  filter(Q29 != "[DNRO] Refused") %>%
  # get proportions
  group_by(zone) %>%
  mutate(
    Q29_ntotal = sum(Q29_n),
    Q29_p = Q29_n / sum(Q29_n)) %>%
  arrange(zone) ->
prop_Q29

prop_Q29 %>%
  stacked_barchart("Q29",
    "In general, would you say your health is...?")
```

#### Q31: In general, would you say your mental health is...?

```{r}
#| label: mentalhealth-general
survey_merged %>%
  select(zone, Q31) %>%
  mutate(Q31 = as_factor(Q31)) %>%
  count(zone, Q31, name = "Q31_n") %>%
  # fill the missing responses (due to 0 counts) back in
  complete(Q31, zone, fill = list(Q31_n = 0)) %>%
  filter(Q31 != "[DNRO] Refused") %>%
  # get proportions
  group_by(zone) %>%
  mutate(
    Q31_ntotal = sum(Q31_n),
    Q31_p = Q31_n / sum(Q31_n)) %>%
  arrange(zone) ->
prop_Q31

prop_Q31 %>%
  stacked_barchart("Q31",
    "In general, would you say your mental health is...?")
```


```{r}
prop_Q31 %>%
  filter(Q31_ntotal > 25) %>%
  mutate(zone_label = glue("Zone {zone} (n = {Q31_ntotal})")) %>%
  {
    ggplot(.) +
      aes(x = zone_label, y = Q31_p, fill = Q31) +
      geom_col(position = "stack") +
      coord_flip() +
      scale_fill_brewer(type = "div", palette = "RdBu", direction = -1) +
      labs(
        x = "In general, would you say your mental health is...?"
      )
  }

```

## National responses



## Outstanding questions

Q12/Q13, Q25-Q28: are folks within a climate zone more/less likely to report being comfortable in summer/winter if they have higher rent/less affordable rent/have higher impact with a 10% increase in rent/there's money left over for things

### Warm vs. cool: which do people struggle with more?

```{r}
#| label: summer-winter
list(winterwarm = prop_Q12, summercool = prop_Q13) %>%
  # combine and pivot
  bind_rows(.id = "df") %>%
  rename(Q12_response = Q12, Q13_response = Q13) %>%
  ungroup() %>%
  pivot_longer(
    cols = matches("Q1[23]*"),
    names_to = c("question", ".value"),
    names_sep = "_",
    values_drop_na = TRUE) %>%
  # tidy up season and response
  select(-question) %>%
  rename(question = df) %>%
  arrange(zone) %>%
  filter(response == "No") %>%
  # make ntotal counts uniform across questions (just for exploratory sorting
  # purposes)
  group_by(zone) %>%
  mutate(
    ntotal = min(ntotal),
    signif = if_else(plow[1] <= phigh[2] & plow[2] <= phigh[1], NA_character_, "✅")) %>%
  ungroup() %>%
  # make zone labels, turn %s back into counts, reformat seasons
  mutate(
    nlow = plow * ntotal,
    nhigh = phigh * ntotal,
    season = case_match(question,
      "winterwarm" ~ "Winter",
      "summercool" ~ "Summer")) %>%
  # finally, merge in town names and select
  left_join(town_list, by = "zone", multiple = "all") %>%
  select(zone, town, season, signif, ntotal, n, p, moe, nlow, plow, nhigh, phigh) %>%
  write_csv(here("data", "survey-discomfort-summerwinter.csv")) ->
summer_winter

# merge in town names
  
# visualise
summer_winter
  %>%
  mutate(
    zone_label = glue("{signif} Zone {zone} (n = {ntotal})"),
    zone_label = fct_reorder(zone_label, ntotal, .desc = TRUE)) %>%
  {
    ggplot(.) +
      aes(x = season, y = p, fill = season) +
      geom_col() +
      geom_linerange(aes(ymin = plow, ymax = phigh)) +
      facet_wrap(vars(zone_label)) +
      scale_y_continuous(labels = scales::label_percent()) +
      scale_fill_manual(
        values = c("Summer" = "firebrick", "Winter" = "royalblue"),
        name = "Season") +
      coord_flip() +
      labs(
        title = paste(
          "% renters who aren't be comfortable",
          "in summer heat or winter cold"),
        subtitle = "✅ Statistically significant"
      )
  }
```

Let's widen these results so that we can compare summer and winter on a Flourish map.

```{r}
summer_winter %>%
  filter(!is.na(zone)) %>%
  pivot_wider(names_from = season, values_from = ntotal:phigh) %>%
  mutate(balance = case_when(
    signif == "✅" & p_Summer > p_Winter ~ "More suffering in summer",
    signif == "✅" & p_Summer <= p_Winter ~ "More suffering in winter",
    p_Summer > p_Winter ~ "No significant difference (summer slightly higher)",
    p_Summer <= p_Winter ~ "No significant difference (winter slightly higher)"
  )) %>%
  filter(!is.na(balance)) %>%
  write_csv(here("data", "survey-discomfort-summerwinter-wide.csv")) ->
summer_winter_wide
```

